Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-10-01 files/Chapter6/Validity_Plate.tex
v2 chapter 7.2, started 7.3

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of the Reissner-Mindlin plate model}
In this section, the validity of a cantilever Reissner-Mindlin plate is investigates with reference to a three-dimensional plate model. In the previous section, it was shown that the beam models do not compare well when the parameter $b$ is ``too-large''. A plate model might be a better option.\\

Similarly to how the Timoshenko beam theory improves on the Euler-Bernoulli beam theory, the Reissner-Mindlin plate theory improves on classical plate theory by including the effect of shear stress. The Reisner-Mindlin plate theory is a two-dimensional theory. Therefore it is of interest to investigate the validity of such a plate model.Using \cite{LVV09} as a guide, the validity a cantilever Reisner-Mindlin plate model is investigated.\\

 Figure \ref{fig:plate_sbs} shows the two cantilever plates side by side.
\FloatBarrier
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\caption{Side by side visualization of the cantilever plates.}
			\label{fig:plate_sbs}
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (6.2,-0.1);
					\draw[line width = 0.3mm] (0.8,0.7) -- (7.8,0.7);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (0.8,0.7);
					\draw[line width = 0.3mm] (6.2,-0.1) -- (7.8,0.7);

@@

					\draw[line width = 0.2mm] (0.8,0.7) -- (7.8,0.7);
					\draw[line width = 0.2mm] (-0.8,-0.1) -- (0.8,0.7);
					\draw[line width = 0.2mm] (6.2,-0.1) -- (7.8,0.7);
				\end{tikzpicture}
				\subcaption{Timoshenko Cantilever Beam}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (-0.8,0.1) -- (6.2,0.1);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (6.2,-0.1);
					\draw[line width = 0.3mm] (6.2,-0.1) -- (6.2,0.1);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (-0.8,0.1);

@@

					\draw[line width = 0.1mm,->] (9,1) -- (10,1.6);
					\draw[line width = 0.1mm,->] (9,1) -- (10.4,1);
					\draw[line width = 0.1mm,->] (9,1) -- (9,-0.1);
					\node at (9.2,1.4) {$e_3$};
					\node at (9.7,0.8) {$e_1$};
					\node at (8.8,0.4) {$e_2$};
				\end{tikzpicture}
				\subcaption{Two-Dimensional Cantilever Beam}

@@

Both the Reissner-Mindlin plate and the three-dimensional plate model has a parameter $b$ representing the depth of the plate. The three-dimensional model also have a parameter $h$ representing the width of the plate model.

\subsection{Comparing the mode shapes}
\FloatBarrier
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Plate2.png}
				\subcaption{Plate - $\lambda_5 = 0.643$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D2.png}
				\subcaption{3D Model - $\lambda_5 = 0.645$}
				\label{fig:minipage1}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}{.8\textwidth}
				\centering
				\includegraphics[width=1\linewidth]{Plate1.png}
				\subcaption{Plate - $\lambda_6 = 1.92$}
				\label{fig:minipage1}
			\end{minipage}
			\begin{minipage}{.8\textwidth}
				\centering
				\includegraphics[width=1\linewidth]{3D1.png}
				\subcaption{3D Model - $\lambda_7 = 1.92$}
				\label{fig:minipage2}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Plate3.png}
				\subcaption{Plate - $\lambda_8 = 2.74$}
				\label{fig:minipage1}
			\end{minipage}
			\quad
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D3.png}
				\subcaption{3D Model - $\lambda_8 = 2.74$}
				\label{fig:minipage2}
			\end{minipage}
	}}
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{Plate4.png}
				\subcaption{Plate - $\lambda_{12} = 8.96$ (Side view)}
				\label{fig:minipage2}
			\end{minipage}
			\quad
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{3D4.png}
				\subcaption{3D Model - $\lambda_{14} = 9.01$ (Side view)}
				\label{fig:minipage2}
			\end{minipage}
			\caption{Comparison of Mode Shapes of the 3D model and the plate model. $b = 1/20$, $d = 1$.}
	}}
\end{figure}

\FloatBarrier

Let $b= 1$ and $\alpha= 4800$ so that $h= \frac{1}{2}$.\\

\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}{.8\textwidth}
				\centering
				\includegraphics[width=1\linewidth]{3Dnp1.png}
				\subcaption{3D Model - $\lambda_{6} = 1.35$ (Top view)}
				\label{fig:minipage1}
			\end{minipage}
			\begin{minipage}{.8\textwidth}
				\centering
				\includegraphics[width=1\linewidth]{3Dnp2.png}
				\subcaption{3D Model - $\lambda_{13} = 7.80$ (Top view)}
				\label{fig:minipage2}
			\end{minipage}
	}}
	\caption{Examples of non-plate mode shapes.}
\end{figure}

\subsection{Comparing the eigenvalues}
Tables \ref{tab:Table_plate_1} and \ref{tab:Table_plate_1} compare the eigenvalues of a cantilever Reissner-Mindlin plate model to the eigenvalues of a cantilever three-dimensional plate model. The different widths of plates considered are $b = 0.25$ and $b = 1$. These cases represent a slightly and very wide plate models. In each of the tables, the parameter $h$ representing the height is decreased.

\FloatBarrier
\begin{table}[htbp]
	\scalebox{.8}{
	\makebox[\textwidth]{
		\caption{Comparison of eigenvalues of the cantilever Reissner-Mindlin and three-dimensional plates.}
		\begin{tabular}{|cccc||cccc||cccc||cccc|}
			\hline
			\multicolumn{16}{|c|}{Comparison of Eigenvalues, $b = 0.25$} \\
			\hline\hline
			\multicolumn{4}{|c||}{$h = 1/5$}       & \multicolumn{4}{c||}{$h =1/10$}      & \multicolumn{4}{c||}{$h = 1/20$}      & \multicolumn{4}{c|}{$h = 1/30$} \\
			\hline
			{i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} \\
			\hline
			1     & 0.12348 & 1     & {0.12249} & 1     & 0.032327 & 1     & 0.032184 & 1     & 0.008207 & 1     & 0.008186 & 1     & 0.003663 & 1     & 0.003656 \\
			2     & 0.18638 &       & -     & 2     & 0.18531 &       & {-} & 2     & 0.18476 &       & {-} & 2     & 0.14204 & 2     & 0.14173 \\
			3     & 2.4151 & 2     & {2.3954} & 3     & 1.161 & 2     & 1.1537 & 3     & 0.31436 & 2     & 0.31341 & 3     & 0.18456 &       & {-} \\
			4     & 3.5856 & 3     & {3.5368} & 4     & 1.3155 & 3     & 1.31  & 4     & 0.44696 & 3     & 0.44582 & 4     & 0.21514 & 3     & 0.21454 \\
			5     & 4.785 &       & -     & 5     & 4.7697 &       & {-} & 5     & 2.3862 & 4     & 2.3767 & 5     & 1.1001 & 4     & 1.0971 \\
			6     & 7.7889 &       & -     & 6     & 7.7674 & 4     & 7.9816 & 6     & 4.1984 & 5     & 4.1857 & 6     & 2.0374 & 5     & 2.0313 \\
			7     & 20.37 & 4     & {19.969} & 7     & 8.056 &       & {-} & 7     & 4.7615 &       & {-} & 7     & 4.1509 & 6     & 4.1367 \\
			8     & 21.722 & 5     & {21.537} & 8     & 12.034 & 5     & 11.972 & 8     & 7.7543 &       & {-} & 8     & 4.7585 &       & {-} \\
			9     & 25.181 &       & -     & 9     & 25.147 &       & {-} & 9     & 8.7609 & 6     & 8.7132 & 9     & 6.2078 & 7     & 6.1864 \\
			10    & 56.321 & 6     & {54.887} & 10    & 26.608 & 6     & 26.266 & 10    & 12.598 & 7     & 12.548 & 10    & 7.7495 &       & {-} \\
			11    & 60.267 & 7     & {59.71} & 11    & 34.438 & 7     & 34.197 & 11    & 22.619 & 8     & 22.456 & 11    & 11.064 & 8     & 11.015 \\
			12    & 65.513 &       & -     & 12    & 61.829 & 8     & 60.801 & 12    & 25.128 &       & {-} & 12    & 13.734 & 9     & 13.678 \\
			13    & 69.031 &       & -     & 13    & 65.518 &       & {-} & 13    & 27.297 & 9     & 27.155 & 13    & 23.854 & 10    & 23.724 \\
			14    & 114.09 & 8     & {110.7} & 14    & 69.133 & 9     & 69.548 & 14    & 47.123 & 10    & 46.695 & 14    & 25.121 &       & {-} \\
			15    & 117.88 & 9     & {116.66} & 15    & 70.211 &       & {-} & 15    & 50.515 & 11    & 50.174 & 15    & 26.054 & 11    & 25.927 \\
			16    & 127.05 &       & -     & 16    & 116.78 & 10    & 114.43 & 16    & 65.523 &       & {-} & 16    & 36.259 & 12    & 36.15 \\
			17    & 184.46 & 10    & {185.78} & 17    & 121.43 & 11    & 119.94 & 17    & 69.089 &       & {-} & 17    & 41.973 & 13    & 41.807 \\
			18    & 191.91 & 11    & {191.47} & 18    & 127.23 &       & {-} & 18    & 71.563 & 12    & 71.166 & 18    & 44.95 & 14    & 44.685 \\
			19    & 193.03 &       & -     & 19    & 178.27 & 12    & 175.79 & 19    & 81.444 & 13    & 80.894 & 19    & 45.798 & 15    & 45.507 \\
			20    & 193.71 &       & -     & 20    & 186.06 & 13    & 187.28 & 20    & 84.785 & 14    & 84.064 & 20    & 54.838 & 16    & 54.573 \\
			\hline
			\multicolumn{2}{|c}{Max RE:} & \multicolumn{2}{c||}{2.9764\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{2.7585\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{0.90845\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c|}{0.63525\%} \\
			\hline
		\end{tabular}%
		\label{tab:Table_plate_1}%
	}}
\end{table}%
\FloatBarrier

The results in Table \ref{tab:Table_plate_1} show that the cantilever Reissner-Mindlin plate model compare very well to the cantilever three-dimensional plate. For $h = 1/5$ the plate can be considered ``thick'', and the maximum relative error is less than $3\%$. This relative error decreases significantly as $h$ decreases.\\

As $b$ decreases, there are also more eigenvalues of the cantilever the Reissner-Mindlin plate relating to the first 20 eigenvalues of the cantilever three-dimensional plate. Table
\ref{tab:Table_plate_1_1} shows some more results for different values of $b$ without presenting the eigenvalues.

\FloatBarrier
\begin{table}[htbp]
	\centering
	\caption{Maximum relative error of the first 10 comparable eigenvalues with b = 0.25.}
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		\multicolumn{6}{|c|}{Max Relative Error } \\ 
		\hline 
		\hline
		b = 1/2 & b = 1/5 & b = 1/10 & b = 1/20 & b = 1/30 & b = 1/50\\ \hline
		3.4514\% & 2.9764 \% & 2.7585 \% & 0.90845 \% & 0.63525 \% & 0.50397 \% \\ \hline
	\end{tabular}%
	\label{tab:Table_plate_1_1}%
\end{table}%

\FloatBarrier
\begin{table}[htbp]
	\scalebox{.8}{
	\makebox[\textwidth]{
		\caption{Comparison of eigenvalues of the cantilever Reissner-Mindlin and three-dimensional plates.}
		\begin{tabular}{|cccc||cccc||cccc||cccc|}
			\hline
			\multicolumn{16}{|c|}{Comparison of Eigenvalues, $b = 1$} \\
			\hline\hline
			\multicolumn{4}{|c||}{$h = 1/5$}       & \multicolumn{4}{c||}{$h =1/10$}      & \multicolumn{4}{c||}{$h = 1/20$}      & \multicolumn{4}{c|}{$h = 1/30$} \\
			\hline
			{i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} & {i} & {3D} & {j} & {Plate} \\
			\hline
			1     & 0.12869 & 1     & {0.12743} & 1     & 0.033784 & 1     & 0.033626 & 1     & 0.008565 & 1     & 0.008543 & 1     & 0.003817 & 1     & {0.00381} \\
			2     & 0.62189 & 2     & {0.61703} & 2     & 0.18635 & 2     & 0.18562 & 2     & 0.049829 & 2     & 0.04957 & 2     & 0.022511 & 2     & {0.022405} \\
			3     & 1.3638 &       & -     & 3     & 1.1603 & 3     & 1.153 & 3     & 0.31428 & 3     & 0.31315 & 3     & 0.14192 & 3     & {0.14156} \\
			4     & 3.5808 & 3     & {3.5289} & 4     & 1.3579 &       & {-} & 4     & 0.50842 & 4     & 0.50676 & 4     & 0.23035 & 4     & {0.22969} \\
			5     & 5.8165 & 4     & {5.7694} & 5     & 1.864 & 4     & 1.8577 & 5     & 0.64595 & 5     & 0.64297 & 5     & 0.29502 & 5     & {0.29394} \\
			6     & 6.6048 & 5     & {6.5216} & 6     & 2.2928 & 5     & 2.2792 & 6     & 1.3555 &       & {-} & 6     & 0.8928 & 6     & {0.88735} \\
			7     & 7.8455 &       & -     & 7     & 6.4973 & 6     & 6.4546 & 7     & 1.9293 & 6     & 1.9157 & 7     & 1.156 & 7     & {1.1527} \\
			8     & 9.8027 &       & -     & 8     & 7.8167 &       & {-} & 8     & 2.5089 & 7     & 2.4975 & 8     & 1.265 & 8     & {1.261} \\
			9     & 17.03 & 6     & {16.802} & 9     & 8.4533 & 7     & 8.3671 & 9     & 2.7486 & 8     & 2.7369 & 9     & 1.3543 &       & - \\
			10    & 21.225 & 7     & {20.743} & 10    & 9.3453 & 8     & 9.2873 & 10    & 3.3087 & 9     & 3.2903 & 10    & 1.5352 & 9     & {1.529} \\
			11    & 23.935 & 8     & {23.552} & 11    & 9.8009 &       & {-} & 11    & 5.5365 & 10    & 5.4914 & 11    & 2.598 & 10    & {2.5803} \\
			12    & 24.694 &       & -     & 12    & 10.925 & 9     & 10.827 & 12    & 5.9696 & 11    & 5.9246 & 12    & 2.8161 & 11    & {2.7997} \\
			13    & 27.187 & 9     & {26.69} & 13    & 17.599 & 10    & 17.447 & 13    & 7.8024 &       & {-} & 13    & 4.2654 & 12    & {4.2474} \\
			14    & 28.844 &       & -     & 14    & 18.596 & 11    & 18.407 & 14    & 9.0173 & 12    & 8.9573 & 14    & 4.6674 & 13    & {4.6415} \\
			15    & 32.373 &       & -     & 15    & 24.729 &       & {-} & 15    & 9.8006 & 13    & 9.838 & 15    & 4.9397 & 14    & {4.9143} \\
			16    & 41.288 & 10    & {40.519} & 16    & 27.4  & 12    & 27.021 & 16    & 9.9172 &       & {-} & 16    & 5.7243 & 15    & {5.6803} \\
			17    & 42.082 & 11    & {41.246} & 17    & 28.81 &       & {-} & 17    & 10.365 & 13    & 10.286 & 17    & 6.6585 & 16    & {6.6025} \\
			18    & 51.427 &       & -     & 18    & 30.737 & 13    & 30.413 & 18    & 11.934 & 14    & 11.817 & 18    & 7.2935 & 17    & {7.2474} \\
			19    & 56.963 & 12    & {56.1} & 19    & 30.83 & 14    & 30.413 & 19    & 13.915 & 15    & 13.767 & 19    & 7.7966 &       & - \\
			20    & 57.741 &       & -     & 20    & 32.414 &       &       & 20    & 15.109 & 16    & 14.974 & 20    & 9.7996 &       & - \\
			\hline
			\multicolumn{2}{|c}{Max RE:} & \multicolumn{2}{c||}{2.2695\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{1.3816\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c||}{1.0657\%} & \multicolumn{2}{c}{Max RE:} & \multicolumn{2}{c|}{0.84102\%} \\
			\hline
		\end{tabular}%
		\label{tab:Table_plate_2}%
	}}
\end{table}%
\FloatBarrier

Similarly, Table \ref{tab:Table_plate_2} also show that a cantilever Reissner-Mindlin plate compares very well to a cantilever three-dimensional plate for $b=1$ which can be considered a wide plate.\\

Table \ref{tab:Table_plate_2_2} presents some more results without showing all the eigenvalues.

\FloatBarrier
\begin{table}[htbp]
	\centering
	\caption{Maximum relative error of the first 10 comparable eigenvalues with d=1.}
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		\multicolumn{6}{|c|}{Max Relative Error } \\ 
		\hline 
		\hline
		b = 1/2 & b = 1/5 & b = 1/10 & b = 1/20 & b = 1/30 & b = 1/50\\ \hline
		3.9342\% & 2.2695 \% & 1.3816 \% & 1.0657 \% & 0.84101 \% & 0.38181 \% \\ \hline
	\end{tabular}%
	\label{tab:addlabel}%
\end{table}%

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of a Reissner-Mindlin plate} \label{sec:validity-of-a-plate}

In this section, the validity of a cantilever Reisner-Mindlin plate model is investigated. The idea of the model comes from the beam models in the previous section failing to approximate a cantilever elastic body with a width much larger than it's height. The process of determining the validity of the model follows the same ideas and methodology as in the previous sections, that is established in \cite{LVV09}.

Similarly to how the Timoshenko beam theory improves on the Euler-Bernoulli beam theory by including the shear forces, so does the Reissner-Mindlin plate models improve on classic plate models. The Reissner-Mindlin plate model is a two-dimensional model, and therefore is it of value to investigate the validity of the model using a three-dimensional plate model as a reference.

\subsection{The models}
Form section \ref{ssec:P_Model:ModelProblems}, the cantilever Reissner-Mindlin plate model is given, and is referred to as Problem P-1. The three-dimensional model is the same as used in the previous section, and is referred to as Problem 3D-1 defined in section \ref{ssec:3D_Model:ModelProblems}.

Figure \ref{fig:plate_sbs} shows the two cantilever plates side by side.

\FloatBarrier
\begin{figure}[h!]
	\scalebox{.7}{
		\makebox[\textwidth][c]{
			\caption{Side by side visualization of the cantilever plates.}
			\label{fig:plate_sbs}
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (6.2,-0.1);
					\draw[line width = 0.3mm] (0.8,0.7) -- (7.8,0.7);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (0.8,0.7);
					\draw[line width = 0.3mm] (6.2,-0.1) -- (7.8,0.7);

@@

					\draw[line width = 0.2mm] (0.8,0.7) -- (7.8,0.7);
					\draw[line width = 0.2mm] (-0.8,-0.1) -- (0.8,0.7);
					\draw[line width = 0.2mm] (6.2,-0.1) -- (7.8,0.7);
				\end{tikzpicture}
				\subcaption{Reissner-Mindlin cantilever plate}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\begin{tikzpicture}
					\draw[line width = 0.3mm] (-0.8,0.1) -- (6.2,0.1);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (6.2,-0.1);
					\draw[line width = 0.3mm] (6.2,-0.1) -- (6.2,0.1);
					\draw[line width = 0.3mm] (-0.8,-0.1) -- (-0.8,0.1);

@@

					% \draw[line width = 0.1mm,->] (9,1) -- (10,1.6);
					% \draw[line width = 0.1mm,->] (9,1) -- (10.4,1);
					% \draw[line width = 0.1mm,->] (9,1) -- (9,-0.1);
					% \node at (9.2,1.4) {$e_3$};
					% \node at (9.7,0.8) {$e_1$};
					% \node at (8.8,0.4) {$e_2$};
				\end{tikzpicture}
				\subcaption{Three-dimensional cantilever plate}

@@

\subsection{Calculating the eigenvalues}
The eigenvalues for both models are calculated using the Finite Element Method. For the plate model, the Finite Element Method is derived in \ref{sec:FEM:Plate} and for the three-dimensional model, the Finite Element Method is derived in \ref{sec:FEM:3D}.

\subsubsection{Problem P-1E}
Find a vector function $u$ and a number $\lambda$ such that
\begin{eqnarray}
	K{u} & = & M\lambda{u}, 
\end{eqnarray} where $K$ and $M$ are the standard Finite Element Method matrices for bi-cubic basis functions. The definition of the matrices are given in \ref{sec:FEM:Plate}.

\subsubsection{Problem 3D-1E}
Find a real number $\lambda$ and a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	K\bar{u} & = & M\lambda{\bar{u}},
\end{eqnarray} where $K$ and $M$ are the standard Finite Element Method matrices for tri-cubic basis functions. The definition of the matrices are given in section \ref{sec:FEM:3D}.

\subsubsection{Accuracy of the eigenvalues}
Figure \ref{fig:conv_3d_eig} show the rate of convergence of the first 20 eigenvalues of Problem 3D-1E.

\begin{figure}[H]
    \centering
    \begin{adjustbox}{center}
        \includegraphics[scale=0.7]{Convergence.png}
    \end{adjustbox}
    \caption{Rate of convergence of the first 20 eigenvalues.}
    \label{fig:conv_3d_eig}
\end{figure}
\textcolor{red}{Update this figure.}

The number of elements can be chosen so that at least the first 10 eigenvalues are accurate to 5 significant digits.

\subsection{Comparing the mode shapes}

\subsection{Comparing the eigenvalues}
